---
title: "MA213 Basic Statistics and Probability — Lab 7 Guide"
subtitle: "Simple Linear Regression"
bibliography: references.bib
output:
  html_document:
    code_download: true
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
```

------------------------------------------------------------------------

## Learning Objectives

By the end of this lab you will be able to:

-   **Compute and Interpret Correlation and R²**: Compute and interpret correlation coefficients and R² values, while recognizing that correlation does not imply causation.
-   **Describe and Assess Relationships Between Two Variables**: Describe the association between two numerical variables in a scatter plot in terms of direction, shape (linear or nonlinear), and strength, and assess whether linear regression is an appropriate model.
-   **Fit and Interpret Linear Models Using Least Squares**: Fit the intercept and slope of a linear model to data using the least squares method, interpret the fitted values, and use the model to predict responses to new inputs.
-   **Perform Inference for Regression Coefficients**: Use fit summary and parameter estimates (e.g., $\hat{\beta_1}$ and $\hat{\sigma^2}$ ) to perform hypothesis tests or construct confidence intervals for the slope, and interpret the results.

------------------------------------------------------------------------

## Background: Beauty in the Classroom

Many college courses end with students anonymously evaluating the instructor. However, these evaluations are often criticized for reflecting non-teaching factors — including the instructor's physical appearance. Researchers at UT Austin collected data on teaching evaluation scores and standardized beauty scores for 463 professors [@hamermesh2005beauty].

> **Beauty score:** 0 = average; negative = below average; positive = above average.\
> **Eval score:** 1–5 scale; higher is better.

**Dataset variables:**

| Variable      | Description                                            |
|---------------|--------------------------------------------------------|
| `minority`    | Is the professor from a non-Caucasian ethnic minority? |
| `age`         | Professor's age                                        |
| `gender`      | Professor's gender                                     |
| `credits`     | `"single"` (elective) or `"more"` (academic course)    |
| `beauty`      | Average attractiveness rating, centered at 0           |
| `eval`        | Average teaching evaluation score (1–5)                |
| `division`    | Upper or lower division course                         |
| `native`      | Is the professor a native English speaker?             |
| `tenure`      | Tenured/tenure-track, or not                           |
| `students`    | Number of students who participated in the evaluation  |
| `allstudents` | Number of students enrolled in the course              |
| `prof`        | Unique numerical identifier for the professor          |

------------------------------------------------------------------------

## Step 1 — Import and Explore the Data

We load `beauty.csv` and take a first look at its structure.

```{r step1-import}
df <- read.csv("beauty.csv")

glimpse(df)
head(df)
```

**Response and explanatory variables:**\
We treat `eval` as the **response** ($Y$) and `beauty` as the **explanatory** variable ($X$). The research question is: does a professor's appearance predict how students rate their teaching?

Here are summary statistics for the two variables of interest:

| Statistic | `beauty` | `eval` |
|-----------|----------|--------|
| Min       | −1.450   | 2.100  |
| Median    | −0.068   | 4.000  |
| Mean      | 0.000    | 3.998  |
| Max       | 1.970    | 5.000  |
| Std. Dev. | 0.788    | 0.554  |

: Descriptive statistics for `beauty` and `eval` (from `summary()`)

------------------------------------------------------------------------

## Step 2 — Visualize and Measure the Relationship

### Scatterplot

A scatterplot lets us assess direction, form, and strength before fitting any model.

```{r step2-scatter}
ggplot(data = df, aes(x = beauty, y = eval)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.8) +
  labs(
    title = "Teaching Evaluation vs. Beauty Score",
    x = "Beauty Score (standardized)",
    y = "Teaching Evaluation (1–5)"
  ) +
  theme_minimal()
```

### Correlation Coefficient

The sample correlation coefficient $R$ measures the **strength and direction** of the linear association:

$$
R = \frac{1}{n-1} \sum_{i=1}^{n} \frac{x_i - \bar{x}}{s_x} \cdot \frac{y_i - \bar{y}}{s_y}
$$

where $s_x$ and $s_y$ are the standard deviations of $X$ and $Y$.

```{r step2-correlation}
n    <- nrow(df)
X    <- df$beauty
Y    <- df$eval

xbar <- mean(X);  ybar <- mean(Y)
sx   <- sd(X);    sy   <- sd(Y)

# Manual calculation (matches the formula above)
R <- sum(((X - xbar) / sx) * ((Y - ybar) / sy)) / (n - 1)
R

# Verify with built-in function
cor(X, Y)
```

**Interpretation:**\
-\> There is a **weak positive** linear association between beauty score and teaching evaluation ($R \approx 0.187$). Beauty explains only a small portion of the variability in course ratings.

------------------------------------------------------------------------

## Step 3 — Fit the Least Squares Line

The simple linear regression model is:

$$y = \beta_0 + \beta_1 x + \varepsilon$$

The least squares estimates minimize the sum of squared residuals. A convenient formula using $R$, $s_x$, and $s_y$ is:

$$\hat\beta_1 = \left(\frac{s_y}{s_x}\right) R, \qquad \hat\beta_0 = \bar{y} - \hat\beta_1\, \bar{x}$$

```{r step3-coefficients}
b1 <- (sy / sx) * R
b0 <- ybar - b1 * xbar

cat("Slope     b1 =", round(b1, 4), "\n")
cat("Intercept b0 =", round(b0, 4), "\n")
```

**Interpretation:**

-   $\hat\beta_1 \approx 0.133$: For each one-unit increase in beauty score, the predicted teaching evaluation increases by about 0.133 points, on average.
-   $\hat\beta_0 \approx 3.998$: A professor with an average beauty score (0) is predicted to receive an evaluation of about 4.0.

------------------------------------------------------------------------

## Step 4 — Compute Fitted Values and Residuals

The **fitted value** for observation $i$ is $\hat y_i = \hat\beta_0 + \hat\beta_1 x_i$.\
The **residual** is $e_i = y_i - \hat y_i$ — the leftover variation not explained by the model.

```{r step4-residuals}
yhat      <- b0 + b1 * X
residuals <- Y - yhat

# Quick check: residuals should sum to (approximately) zero
sum(residuals)
```

------------------------------------------------------------------------

## Step 5 — Compute $R^2$

$R^2$ measures the **proportion of variability in** $Y$ explained by $X$:

$$R^2 = R \times R$$

```{r step6-r2}
R_squared <- R^2
cat("R-squared =", round(R_squared, 4), "\n")
```

**Interpretation:**\
-\> About **3.5%** of the variation in teaching evaluations is explained by beauty score. The model fit is poor — most of the variation in `eval` is due to factors other than appearance.

------------------------------------------------------------------------

## Step 7 — Using `lm()` and Reading the Summary

R's `lm()` function fits the same model in one line. Its `summary()` output bundles everything we computed above.

```{r step7-lm}
model         <- lm(Y ~ X)
model_summary <- summary(model)
model_summary
```

### Extracting Key Quantities

```{r step7-extract}
# Coefficients (b0 and b1)
model_summary$coefficients

# First six residuals
head(model_summary$residuals)

# R-squared
model_summary$r.squared

# Adjusted R-squared (penalizes for number of predictors)
model_summary$adj.r.squared
```

**Do the `lm()` values match our manual calculations?**

```{r step7-compare}
cat("Manual  b0:", round(b0, 6), "  lm b0:", round(coef(model)[1], 6), "\n")
cat("Manual  b1:", round(b1, 6), "  lm b1:", round(coef(model)[2], 6), "\n")
cat("Manual  R²:", round(R_squared, 6), "  lm R²:", round(model_summary$r.squared, 6), "\n")
```

------------------------------------------------------------------------
