%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Lesson Plan (50 minutes)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
    \frametitle{Lesson Plan}
    \begin{itemize}
        \item xx min Lecture: reminder of two-sample power with figure. What happens when the effect size is small?
        \item xx min R demo: Shiny app (lecture 27) to show what happens when the effect size is small
        \item xx min Edfinity quiz: picture of a very low power case. What is true? 
        \item xx min Lecture: Introduction to replication crisis, highlight Ioannidis paper
        \item xx min Lecture: retraction watch quote on Kahneman "placed too much faith"
        \item xx min Lecture: Beauty and Sex Ratios, compute standard error
        \item xx min Think/pair/share: Show the power figure and ask
        \begin{enumerate}
            \item Approximately how big did the effect size have to be to reach significance?
            \item Approximately what was the power of the test to detect an effect of 1\%?
            \item What is the approximate probability that a statistically significant result is in the wrong direction (sign error)?
        \end{enumerate}
        \item xx min Lecture: Results
        \item xx min Lecture: How many subjects would be required to detect an effect of 1\%? 
        \item xx min Lecture: Takeaways and best practices
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Learning objectives:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
    \frametitle{Learning Objectives}
    \begin{itemize}
        \item \textbf{M1, LO3: Use R for Data Management and Exploration:} Utilize R to load, pre-process, and explore data through visualization and summarization techniques.
        \item \textbf{M3, LO1: Understand Point Estimates and Sampling Variability:} Define a sample statistic (point estimate) for a population parameter, and explain how it varies across different samples.
        \item \textbf{M3, LO3: Calculate and Interpret Standard Error:} Calculate the standard error for proportions and interpret it as a measure of sampling variability.
        \item \textbf{M3, LO4: Explain Hypothesis Testing and Its Limitations:} Discuss the use cases and potential issues with hypothesis testing, including the interpretation of results.
        \item \textbf{M3, LO6: Distinguish Statistical vs. Practical Significance:} Differentiate between statistical significance and practical significance, and explain the implications of each.
    \end{itemize}
\end{frame}
    
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: Edfinity quiz: HT concepts

