---
title: "MA213 - Fall 2025"
output:
  pdf_document: default
  html_document:
    code_download: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# **Lab Session 4: Simulation and probability**

## Basic Statistics and Probability

In this lab session, we will go over central limit theorem. Can we verify central limit theorem by simulation?

## Pre-lab activities

Let's think about

-   How do we get random sample from a specific distribution?

-   Can we visualize the shape of the distribution you picked?

## Lab activities

1.  You are given different population. Each group pick different population. We are supposed to pick it up differently.

#### Some distribution functions in `R`.

-   **Normal distribution** : `rnorm(n, mean, sd)`

    -   `n`: Number of random values to generate
    -   `mean`: Mean (center) of the distribution
    -   `sd`: Standard deviation (spread) of the distribution

-   **Exponential distribution** : `rexp(n, rate)`

    -   `n`: Number of random values to generate
    -   `rate`: Rate parameter (inverse of the mean)
    -   Mean is `1/rate`
    -   Standard deviation: `1/rate`

-   **Binomial distribution** : `rbinom(n, size, prob)`

    -   `n`: Number of random values to generate
    -   `size`: Number of trials
    -   `prob`: Probability of success in each trial
    -   Mean is `n*prob`
    -   Standard deviation is `sqrt(size * prob * (1 - prob))`

-   **Poisson distribution** : `rpois(n, lambda)`

    -   `n`: Number of random values to generate
    -   `lambda`: Expected number of occurrences in a fixed interval
    -   Mean is `lambda`
    -   Standard deviation is `sqrt(lambda)`

-   **Gamma distribution** : `rgamma(n, shape, scale)`

    -   `n`: Number of random values to generate
    -   `shape`: Shape parameter (determines the distribution's form)
    -   `scale`: Scale parameter (stretches or shrinks the distribution)
    -   The mean is `shape*scale`
    -   Standard deviation is `sqrt(shape * scale^2)`

-   **Uniform distribution** : `runif(n, min, max)`

    -   `n`: Number of random values to generate
    -   `min`: Lower bound of the distribution
    -   `max`: Upper bound of the distribution
    -   Mean is `(min + max) / 2`
    -   Standard deviation is `sqrt((max - min)^2 / 12)`

```{r}
sigma <- 0.5
N <- 1000000

# Poisson Distribution (lambda = sigma^2)
lambda <- sigma^2
pops <- rpois(N, lambda = lambda)


```

What is the shape of the data?

```{r}


# histogram
hist(pops, probability = TRUE)
lines(density(pops), col = "red", lwd = 2)
```

What is mean and standard deviation of the population?

```{r}
mean(pops)
sd(pops)
```

Let's obtain $n=10$ and obtain $\bar{X}$ for 100 times from each sample.

```{r}


```

Please also show the shape of the $\bar{X}$

```{r}



```

2.  Show some of the obtained $\bar{X}_{i}$. Obtain the mean and the standard deviation of the $\bar{X}$.

```{r}


```

3.  What is your conclusion?

How can you tell the distribution follows normal distribution?

```{r}


```

4.  Can we do when n=100 ? What do you see?

```{r}


```

### Post-lab activities

Recent statistics suggest that 10% of Americans rely on Social Network Sites (SNSs). We are going to have a simulation study if we can verify that CLT really works in this case too. Let's assume that US population is 300,000 for calculation purposes.

Please name your submission as `lab4.R`

0.  execute this `set.seed(213)`

1.  Produce population data and assign it `population` object.

2.  Obtain mean and standard deviation of the population and assign them to `mean_pop` and `sd_pop` accordingly.

3.  Get the sample mean data for $n=10$ with $k=100$ iterations. Assign the mean and standard deviation of the sample mean data to `mean_sample` and `sd_sample`.

4.  Do one more time with the sample mean data for $n=50$ with $k=100$ iterations. Assign the mean and standard deviation of the sample mean data to `mean_sample2` and `sd_sample2`.

### Extra question?

Suppose a random sample of 200 people are taken.

What is the distribution of the sample proportion? Explain.

What is the mean and standard deviation of the sample proportion?

What is the probability that less than 32 of the people in the sample rely on SNSs?

What is the probability that more than 20% of the people in the sample rely on SNSs?

What is the probability that between 100 and 140 of the people in the sample rely on SNSs?
