---
title: "MA213 Basic Statistics and Probability - Lab3 Activity"
bibliography: references.bib
link-citations: true
output:
  html_document:
    code_download: true 
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## **Lab 3:** **Plotting and Summaries** activities

------------------------------------------------------------------------

#### Q1. Import the `airquality.csv` file and assign it do `df1`. Discuss with your partner the following questions.

1.  What kind of exploratory data analysis (EDA) can be done on this dataset?
2.  What could be a research question or interest based on this data?

```{r}
#
#
# 
```

#### Q2. Import the Titanic dataset and assign it to `df2`. Discuss with your partner the following questions.

1.  What kind of exploratory data analysis (EDA) can be done on this dataset?
2.  What could be a research question or interest based on this data?

```{r}


```

#### Q3. Show the bar plot of average Solar radiation per each month in the Airquality data. What do you see from the plot?

```{r}
#
#
#

```

#### Q4. Create a contingency table to examine the relationship between Class and Survived in the Titanic data. Assign it to `my_table1` and print it (Use `dplyr` package functions).

```{r}
#
#
#


```

#### Q5. Make a plot to visualize the relationship between Class and Survived using `ggplot()`.

```{r}
#
#
#

```

and answer these questions with the partner.

1.  Which class had the highest survival count?
2.  What general pattern can you observe between passenger class and survival?

### Post-lab activities

Name your submission as `lab3.R`

We are going to also analyze this New York Air Quality Measurements data again. This time, we are going to see how Solar radiation is behaving with Temperature.

1.  Import the `airquality.csv` file and assign it to `my_df` object.

2.  Create the scatter plot of Solar radiation and Temperature and assign it to `my_plot1` .

3.  Show average Solar radiation for each month using a bar plot.

    3.1) Create a new data frame `new_df` containing the average Solar radiation per month (hint: use summarize()) and name the average column `solar_r_avg`.

    3.2) From `new_df`, create a bar plot of the average Solar radiation for each month and assign it to `my_barplot`.

4.  Import the `titanic.csv` file and assign it to `my_df2` object.

5.  From this `my_df2`, create a contingency table for categorized `Sex` vs `Age` and assign it to `titanic_table` .

6.  How many boy (male and child) are there? assign numeric number to `child_male` .

```{r}

# You may write your code here. Copy the code into `lab3.R` and submit it to Gradescope.

# Do not use setwd() here.

library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)

#1


#2


#3.1


#3.2


#4


#5


#6



```
