\documentclass[12pt]{article}
\usepackage{resource/shortex}
\input{resource/teaching}

%--------------------
\begin{document}

\begin{center}
\textbf{\Large Deliverable 1 \\ Project 1 Proposal Outline} 
\end{center}

% \section*{Project Proposal Sections}
\subsection*{Overview}
You have a project deliverable due Next Friday at 8:00 AM. It should be a 1-2 pages outline of your project.
This outline should include the following sections:


\subsection*{1. Project Title}
\begin{itemize}
    \item Provide a concise and descriptive project title.
\end{itemize}
% \vspace{4\baselineskip}

\subsection*{2. Group Members}
\begin{itemize}
    \item List the names and student IDs for all group members.
\end{itemize}
% \vspace{4\baselineskip}

\subsection*{3. Background and Motivation}
\begin{itemize}
    \item Briefly describe the context and motivation for your project.
    \item Explain why your group has chosen this topic or dataset.
\end{itemize}
% \vspace{4\baselineskip}

\subsection*{4. Research Question(s) / Objectives}
\begin{itemize}
    \item Clearly state the primary question(s) your group aims to answer.
    \item List specific objectives or hypotheses (if applicable).
\end{itemize}
% \vspace{4\baselineskip}

\subsection*{5. Data Source and Description}
\begin{itemize}
    \item Describe the data you will use.
    \item Outline the main variables and the size of the dataset.
\end{itemize}
% \vspace{4\baselineskip}

\subsection*{6. Timeline and Responsibilities}
\begin{itemize}
    \item Provide a rough schedule for each major milestone.
    \item Assign tasks or responsibilities to group members.
\end{itemize}
% \vspace{4\baselineskip}


\end{document}