\documentclass[12pt]{article}
\usepackage{resource/shortex}
\input{resource/teaching}


%--------------------
\begin{document}



\begin{center}
\textbf{\Large Project 1 Guide \\ Week 4 } 
\end{center}




\section*{Lab Activities}

\subsection*{1. Let's meet your group and introduce yourselves.  } 

\vspace{6\baselineskip}



\subsection*{2. Discuss about the roles of each group member}

\subsubsection*{Roles and Responsibilities}
Possible roles
\begin{itemize}
    \item \textbf{Facilitator}: Leads discussion with open-ended questions. Summarizes and clarifies group comments. Checks for consensus or questions from group members.
    \item \textbf{Organizer}: Schedules meetings and arranges a timeline/agenda. Takes notes at meetings to send to everyone afterwards. Keeps the meeting on track.
    \item \textbf{Editor(s)}: Compiles and edits different pieces of the project from group members to create consistency.
    \item \textbf{Researcher(s)}: Researches topics and sources for the project and presents the information to the group.
    \item \textbf{Writer(s)}: Writes the final project/report/presentation.
    \item \textbf{Brain-stormer(s)}: Thinks about positives/negatives of ideas presented by the group and identifies possible solutions to problems.
    \item \textbf{Presenter(s)}: Works with group members to compile, create, and present information to the class.
\end{itemize}


\subsection*{3. Discuss about possible topics and data }

\begin{itemize}
    \item Possible numerical data
\end{itemize}

\vspace{2\baselineskip}

\begin{itemize}
    \item Possible categorical data
\end{itemize}

\vspace{2\baselineskip}

\begin{itemize}
    \item Possible questions on thoese data
\end{itemize}

\vspace{2\baselineskip}

\begin{itemize}
    \item Source of data 
\end{itemize}


\vspace{2\baselineskip}

\section*{Post Lab Activities}
You have a project deliverable : the project outline due next Friday at 8:00 AM.






\end{document}
