\documentclass{article}

% --- LaTeX Packages ---
\usepackage[utf8]{inputenc} % For international characters
\usepackage[T1]{fontenc}    % Recommended font encoding
\usepackage{geometry}       % For setting page margins
\geometry{letterpaper, margin=1in} % Set page size and margins

\usepackage{enumitem}       % For customizing lists (useful for instructions/scales)
\usepackage{tabularx}       % For tables with flexible column widths (essential for evaluation criteria)
\usepackage{ragged2e}       % For \RaggedRight in tabularx cells
\usepackage{array}          % Enhanced array and tabular environments

\begin{document}

% Title centered only
\begin{center}
  {\Huge\bfseries Peer Evaluation Form\textsuperscript{*}} % Main title
  \vspace{0.5cm}

  \hrule % Horizontal line
  \vspace{0.5cm}
\end{center}

% Instructions - left aligned for readability
Please write the names of all your team members, INCLUDING YOURSELF, and rate the degree
to which each member fulfilled his/her responsibilities in completing the group assignments.
Use the following rating system:

\begin{itemize}[leftmargin=*, label=\textbullet]
    \item \textbf{Excellent}: Consistently went above and beyond the call of duty, tutored teammates, carried more than his/her fair share of the load.
    \item \textbf{Very good}: Consistently did what he/she was supposed to do, was very well prepared and cooperative.
    \item \textbf{Satisfactory}: Usually did what he/she was supposed to do, was acceptably prepared and cooperative.
    \item \textbf{Ordinary}: Often did what he/she was supposed to do, was minimally prepared and cooperative.
    \item \textbf{Marginal}: Sometimes failed to show up or complete assignments, was rarely prepared.
    \item \textbf{Deficient}: Often failed to show up or complete assignments, was rarely prepared.
    \item \textbf{Unsatisfactory}: Consistently failed to show up or complete assignments, was unprepared.
    \item \textbf{Superficial}: Had practically no participation.
    \item \textbf{No show}: Had no participation at all.
\end{itemize}

\vspace{0.5cm}

These ratings should reflect each individual's level of participation and effort and sense of responsibility, not his or her academic ability.

\vspace{0.5cm}

% Table for names and ratings
\begin{center}
\begin{tabularx}{0.8\textwidth}{>{\RaggedRight\arraybackslash}X|>{\centering\arraybackslash}X}

\textbf{\Large Name of Team Member} & \textbf{\Large Rating} \\
\hline
\vspace{1.5em} & \\ \\
\hline
\vspace{1.5em} & \\ \\
\hline
\vspace{1.5em} & \\ \\
\hline
\vspace{1.5em} & \\ \\
\hline
\vspace{1.5em} & \\ \\
\hline
\end{tabularx}
\end{center}

\vspace{1cm}

\hrule

\vspace{0.2cm}

\noindent * Adapted from Kaufman, D.M., Felder, R.M. and Fuller H (2000). Accounting for Individual Effort in Cooperative Learning Teams. \textit{Journal of Engineering Education}, April 2000, 133–140.

\end{document}