\documentclass{article}
\usepackage[table]{xcolor}
\usepackage{array}

\newcolumntype{P}[1]{>{\raggedright\vrule height4ex width 0pt}p{#1}<{\vrule depth 2.5ex width 0pt}}

\begin{document}

\section*{Project Plan}

\begin{tabular}{|P{2.5cm}*{10}{|c}|}
\hline
\centering \raisebox{-2ex}[0pt][0pt]{Action Plan} & \multicolumn{9}{c|}{Project 1} \\
\cline{2-10}
\multicolumn{1}{|c|}{\vphantom{$\Big|$}} &
\scriptsize Week 1 & \scriptsize Week 2 & \scriptsize Week 3 & \scriptsize Week 4 &
\scriptsize Week 5 & \scriptsize Week 6 & \scriptsize Week 7 & \scriptsize Week 8 & \scriptsize Week 9  \\
\hline
1. Project 1 group announcement &
\multicolumn{2}{c}{\cellcolor{gray}} &&&&&&& \\
\hline
2. Forming groups (modification)&
& \multicolumn{3}{c}{\cellcolor{yellow}} &&&&& \\
\hline
3. Project 1 logistics  &&
&& \multicolumn{1}{c}{\cellcolor{purple}} &&&&& \\
\hline
3. Project 1 work in Lab (week4)  &&
&& \multicolumn{1}{c}{\cellcolor{purple}} &&&&& \\
\hline
3. Deliverable 1 - Outline &
&&& \multicolumn{2}{c}{\cellcolor{blue}} &&&&  \\
\hline
4. Deliverable 2 - Project Progress Report &
&&&& \multicolumn{2}{c}{\cellcolor{blue}} &&&  \\
\hline
5. Deliverable 3 - Presentation Slides &
&&&&& \multicolumn{2}{c}{\cellcolor{blue}} &&  \\
\hline
5. Project 1 work in Lab (week7) &
 &&&&&& \multicolumn{1}{c}{\cellcolor{purple}} && \\
\hline
6. Project 1 Presentation &
 &&&&&&& \multicolumn{1}{c}{\cellcolor{violet}} & \\
\hline
7. Evaluation &
 &&&&&&&& \multicolumn{1}{c}{\cellcolor{cyan}}  \\
\hline
\end{tabular}



\section*{Forming Groups (Project 1)}
\begin{enumerate}
    \item Begin forming groups randomly.
    \item Instructor modifies the groups based on gender and ehtnicity.
    \item Wait until group formation is complete. (2-3 weeks)
    \item Final groups are announced (week 4). 
\end{enumerate}



\section*{Project Deliverables}
\begin{itemize}
    \item Deliverable 1 : Project outline 
    \item Deliverable 2 : Project progress report
    \item Deliverable 3 : Presentation Slides (in PDF format)
\end{itemize}



\subsection*{Peer Evaluation}
After each deliverable, there will be a peer evaluation process.
\begin{itemize}
    \item Excellent, Very good, Satisfactory, Ordinary, Marginal, Deficient, Unsatisfactory, Superficial, No show
    \item If Superficial or No show, the member should conctact the instructor.
    \item If Superficial or No show more than 1, the member will receive unsatisfactory grade for the project.
    \item If Excellent, the member will receive extra credit for the deliverable.
    \item The peer evaluation will be based on the following scale:
    \item Very good(100\%), Satisfactory(90\%), Ordinary(80\%), Marginal(70\%), Deficient(60\%), Unsatisfactory(50\%), Superficial(0\%), No show(0\%).
    \item In order to receive Satisfactory, the member should have average of 65\% or more. 
\end{itemize}



\section*{Roles and Responsibilities}
Possible roles
\begin{itemize}
    \item \textbf{Facilitator}: Leads discussion with open-ended questions. Summarizes and clarifies group comments. Checks for consensus or questions from group members.
    \item \textbf{Organizer}: Schedules meetings and arranges a timeline/agenda. Takes notes at meetings to send to everyone afterwards. Keeps the meeting on track.
    \item \textbf{Editor(s)}: Compiles and edits different pieces of the project from group members to create consistency.
    \item \textbf{Researcher(s)}: Researches topics and sources for the project and presents the information to the group.
    \item \textbf{Writer(s)}: Writes the final project/report/presentation.
    \item \textbf{Brain-stormer(s)}: Thinks about positives/negatives of ideas presented by the group and identifies possible solutions to problems.
    \item \textbf{Presenter(s)}: Works with group members to compile, create, and present information to the class.
\end{itemize}



\section*{Role of TFs and LAs}
\begin{itemize}
    \item TFs and LAs will help with the project.
\end{itemize}



\end{document}