---
title: "MA213 Basic Statistics and Probability — Lab 5 Post-lab Activity"
output:
  html_document:
    code_download: true
    toc: true
    toc_float: true
    theme: flatly
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
```

## Post-Lab Activities

> **Submission instructions:** Save your completed script as `lab5.R` and submit it before Next Tuesday 10pm.

In a hypothetical world, **10% of Americans** rely on Social Network Sites (SNSs) as their primary news source (true). We will conduct a simulation study to verify that the CLT holds in this setting. For computational purposes, assume the U.S. population is $N = 300{,}000$.

**0.** Set the random seed for reproducibility:

``` r
set.seed(213)
```

**1.** Generate the population data and store it in an object called `population`.

**2.** Compute the mean and standard deviation of the population. Assign them to `mean_pop` and `sd_pop`, respectively.

**3.** Simulate the sampling distribution with $n = 10$ and $K = 1{,}000$ iterations. Compute and assign the mean and standard deviation of the resulting sample means to `mean_sample` and `sd_sample`.

**4.** Repeat the above with $n = 50$ and $K = 100$ iterations. Assign the mean and standard deviation of the resulting sample means to `mean_sample2` and `sd_sample2`.

**5.** With $n = 1{,}000$ and $K = 1{,}000$, extract the subsets of sample proportions that fall within 1, 2, and 3 standard deviations of the mean. Assign these subsets to `first_sd_data`, `second_sd_data`, and `third_sd_data`, respectively. Do you think the sample proportions are normally distributed? Answer `normal <- 1` for yes and `normal <- 0` for no.

```{r}

# You may write your code here. Copy the code into `lab5.R` and submit it to Gradescope.

# Do not use setwd() here.

library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)

set.seed(213)

#1


#2


#3



#4



#5




```

## 
