---
title: "MA213 Basic Statistics and Probability - Lab6 Activity"
bibliography: references.bib
output:
  pdf_document: default
  html_document:
    code_download: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## Lab 6: **Confidence Interval and Hypothesis Test**

### Pre-lab activity

### Testing Hypotheses using confidence intervals

Suppose that a market research company is tasked with determining the percentage of adults in Boston who own cell phones. Two hundred randomly selected residents in Boston are surveyed to answer whether they have cell phones. 127 responded yes to the question "do you have a cell phone?"

Recall that

The confidence interval for $p$ is given by $$
\hat{p} \pm (\text{margin of error})
$$

$$
= \hat{p} \pm (z^{*} \times SE)
$$

where $z^{*} = 1.96$ for a 95% CI and

$$
SE = \sqrt{\frac{p(1-p)}{n}   }
$$

We will use the substitution approximation of $p \approx \hat{p}$.

So we use

$$
SE = \sqrt{\frac{\hat{p}(1-\hat{p})}{n} }
$$

#### Exercise 0

Construct 95% confidence interval for the proportion of Boston residents who have cell phones and interpret the confidence interval you found.

```{r}
#
#
#
response <-127
n <- 200
phat <- response/n
alpha <- 0.05

zstar <- qnorm(1-alpha/2, mean = 0, sd = 1)
SE <- sqrt( phat*(1-phat)/n )

CI_lower <- phat - zstar*SE
CI_upper <- phat + zstar*SE

CI <- c(CI_lower, CI_upper)
cat(CI)
```

## Lab activities

Let’s do a simulation using confidence intervals. What we are going to do is repeatedly take random samples from a large “population” in which we know the true proportion who own a cell phone (here, 90%), construct a confidence interval for each sample, and then check whether each interval contains the true population proportion. By doing this many times (e.g., 1000 samples), we can empirically see how often a 95% confidence interval actually includes the true parameter. This lets us see the long‑run meaning of “95% confidence”: in repeated sampling, about 95% of the intervals we construct should capture the true proportion.

### Population

```{r}
N <- 675000
pop <- c(rep("Cellphone", N*0.9), rep("No Cellphone", N*0.1))
true_prop <- sum(pop=="Cellphone")/length(pop)
true_prop


```

Here are steps to see how many confidence intervals contain the true proportion among `K` confidence intervals.

(using alpha level 0.05, that means (1-alpha) % confidence)

1.  Sample `n` from `pop` assign it to `sampled_data`
2.  Obtain sample proportion and assign it to `phat`
3.  Using $SE = \sqrt{ \frac{\hat{p} (1-\hat{p})}{n} }$ , obtain `SE` and obtain a confidence interval
4.  Repeat these steps `K` times and count how many intervals contain the true proportion `true_prop` .

#### Exercise 1. Complete the following R code

---
alpha = 0.05
n <- 200
K <- 1000
num_of_simulation <- rep(0,K)
prop_est <- rep(0,K)
lower <- rep(0,K)
upper <- rep(0,K)
included  <- rep(0,K)

for (k in 1:K) {
  #Step 1
  sampled_data <-   #
    
  #Step 2  
  p_hat <-  #
  
  # Step 3
  SE <-   #
  
  CI = p_hat + c(-1,1)   #
  
  num_of_simulation[k] = k
  prop_est[k] <- p_hat
  lower[k] <- CI[1]
  upper[k] <- CI[2]
  included[k] <- CI[1] <= true_prop & true_prop <=CI[2] 
  
}

# bind together
df_table <- data.frame(num_of_simulation = num_of_simulation, prop_est = prop_est, lower = lower, upper = upper, included = included)

head(df_table)
---

```{r}
# you can copy and paste the completed code above here
#
#



```

```{r}
sum(df_table$included) # number of CIs contains the true population propotion

K # total number of CIs
```

#### Exercise 2. Using the code above and modify it here, change your alpha=0.01 . How many intervals contain the true population parameter?

```{r}




```

```{r}
sum(df_table$included) # number of CIs contains the true population propotion

K # total number of CIs
```

### Testing for independence in two-way tables

We are going to consider hair and eye color of statistics students data. It is the data contains 592 statistics students hair and eye color data.

```{r}
# load the data 
df <- read.csv("haircolor.csv")
head(df)

```

Here are the steps of testing for independence in two-way tables.

1.  Check how many categories each variables have.
2.  Make a contingency table using `table()` function.
3.  Calculate expected counts for each cell.
4.  Make an another contingency table whose cells are expected counts.
5.  Compute the chi-square statistic
6.  Compute p-value and make a decision by comparing the p-value and alpha level.

```{r}

#Step 1 
unique(df$Hair)
unique(df$Eye)



```

```{r}
#Step 2
#Make a contingency table using table() function. 



```

```{r}
# Step 3
# Get the expected counts
# e11, e12, e13, e14, e21,...


```

```{r}
#Step 4
#expected_tbl 

```

```{r}
#Step 5 : Chisquare statistic
#chi_statistic

```

```{r}
#Step 6
#p-value, what is df?
```

What is your conclusion?

## Post-lab activities

Please name your submission as `lab6.R`

1.  The file `a.csv` contains poll results in which random students were asked the question: "Do you own a car?" Assume that these students represent all American college students. Calculate a 95% confidence interval (CI) for the proportion of students who own a car. The sample proportion $\hat{p}$​ can be calculated as the fraction of students who answered "Yes." Obtain `CI` vector.

2.  The file `b.csv` contains Netflix data with variables for plans and gender. At a significance level of $\alpha=0.05$, test whether there is an association between plan type and gender.

    Obtain `df`, `chi_stat` and `p_value` where `df` is the degrees of freedom, `chi_stat` the Chi-square test-statistic, and `p_value` is the p value obtained from the test.

    Based on the test, do you reject $H_0$? If you reject $H_0$, then have a variable `reject_null=T` otherwise, `reject_null=F`

(You need to have variables `df`, `chi_stat`, `p_value` and `reject_null` in your R file)

```{r}
### Warning : do not copy {r} above in the R file ###

##### Lab 6 Post Lab Activity #####

##### Do not install packages in this file #####
##### Do not use the setwd() function in this file #####
##### Do not use the library() function in this file #####

#  The file a.csv contains poll results in which random students were asked the question:
#   "Do you own a car?" Assume that these students represent all American college students. 
#   Calculate a 95% confidence interval (CI) for the proportion of students who own a car. 
#   The sample proportion p_hat can be calculated as the fraction of students who answered "Yes."
#   Obtain CI vector.

data_a <- read.csv("a.csv")





# The file b.csv contains Netflix data with variables for plans and gender. 
# At a significance level of alpha=0.05, 
# test whether there is an association between plan type and gender.
# Obtain df, chi_stat and p_value where df is the degrees of freedom, 
# chi_stat the Chi-square test-statistic, and p_value is the p value obtained from the test.

# Based on the test, do you reject H_0? If you reject H_0, 
# then have a variable reject_null=T otherwise, reject_null=F

data_b <- read.csv("b.csv")


# (You need to have variables df, chi_stat, p_value and reject_null in your R file)
# 
# 
```
