---
title: "MA213 - Fall 2025"
output:
  pdf_document: default
  html_document:
    code_download: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Basic Statistics and Probability

## **Lab Session 2: Plotting and Summaries**

In this lab session, we will go over

## Pre-lab activities

-   Let’s form another group of two. This time, choose a different person.
-   Review of last time and questions from last time.

## Lab activities

Load some libraries

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```

Load the data from `airquality.csv` and show first 5 rows to see how this data is organized.

```{r}


```

Let's check if we have a missing value.

```{r}


```

Let's save them without missing value.

```{r}


```

Let's discuss what is numerical data and what is categorical data. Also, discuss variables of interest (and possible hypothesis).

```{r}


```

|                       |                 |
|-----------------------|----------------:|
| airquality {datasets} | R Documentation |

## New York Air Quality Measurements

### Description

Daily air quality measurements in New York, May to September 1973.

### Usage

```         
airquality
```

### Format

A data frame with 153 observations on 6 variables.

|        |           |         |                         |
|-------:|:----------|:--------|:------------------------|
| `[,1]` | `Ozone`   | numeric | Ozone (ppb)             |
| `[,2]` | `Solar.R` | numeric | Solar R (lang)          |
| `[,3]` | `Wind`    | numeric | Wind (mph)              |
| `[,4]` | `Temp`    | numeric | Temperature (degrees F) |
| `[,5]` | `Month`   | numeric | Month (1--12)           |
| `[,6]` | `Day`     | numeric | Day of month (1--31)    |

### Details

Daily readings of the following air quality values for May 1, 1973 (a Tuesday) to September 30, 1973.

-   `Ozone`: Mean ozone in parts per billion from 1300 to 1500 hours at Roosevelt Island

-   `Solar.R`: Solar radiation in Langleys in the frequency band 4000–7700 Angstroms from 0800 to 1200 hours at Central Park

-   `Wind`: Average wind speed in miles per hour at 0700 and 1000 hours at LaGuardia Airport

-   `Temp`: Maximum daily temperature in degrees Fahrenheit at LaGuardia Airport.

### Source

The data were obtained from the New York State Department of Conservation (ozone data) and the National Weather Service (meteorological data).

### References

Chambers, J. M., Cleveland, W. S., Kleiner, B. and Tukey, P. A. (1983) *Graphical Methods for Data Analysis*. Belmont, CA: Wadsworth.

Numerical Data

mean and standard deviation

1.  Make a scatter plot of `Ozone` and `Temp` .

```{r}



```

2.  Make a dot plot of `Ozone`.

```{r}



```

3.  Make a stacked dot plot of `Ozone` .

```{r}


```

4.  Show histogram of `Ozone` .

```{r}



```

4.  Show histogram of `Ozone` with different bins/breaks to show more information.

```{r}



```

## Categorical Data

We are going to see the famous `Titanic` dataset.

Let's load the data.

```{r}



```

Which one is numerical variable and which one is categorical variable?

How do we know? Let's discuss about them.

```{r}


```

1.  Show the contingency table for `Pclass` vs `Survived` .

```{r}




```

2.Show Bar plot of `Survived`.

```{r}



```

3.  Show the pie chart of `Pclass`.

```{r}


```

4.  We are also interested in the average age of each class. Show the bar plot of average age per each class.

    ```{r}

       


    ```

### Post-lab activities

Please name your submission as `lab2.R`

We are going to also analyze this New York Air Quality Measurements data again. This time, we are going to see how Solar radiation is behaving with Temperature.

0.  load the `airquality.csv` file and assign it to `my_df` object.

1.  Show the scatter plot of Solar radiation and Temperature.

2.  Show average Solar radiation in each month. Use bar plot.

    2.1) Assign average Solar radiation into object `solar_radiation_avg` .

    2.2) Show the bar plot of average Solar radiation per each month.

Back to Titanic data.

0.  Show the pie chart of `sex` in `titanic` data.

    3.1) Load `titanic.csv` file and assign it to `my_df2` object.

    3.2) Show the pie chart of `sex` .
