\documentclass[12pt]{article}
\usepackage{resource/shortex}
\input{resource/teaching}

%--------------------
\begin{document}

\begin{center}
\textbf{\Large Deliverable 1 \\ Final Project Outline} 
\end{center}

\subsection*{Overview}
The final project outline is due Oct 31, Friday at 8:00 AM.  
This outline should be 1--2 pages long and include the following sections to clearly plan your final project.

\subsection*{1. Project Title}
\begin{itemize}
    \item Provide a concise and descriptive title for your project.
\end{itemize}

\subsection*{2. Group Members}
\begin{itemize}
    \item List the full names and student IDs of all group members.
\end{itemize}

\subsection*{3. Background and Motivation}
\begin{itemize}
    \item Briefly describe the context and motivation for your project.  
    \item Explain why your group chose this topic and the relevance or importance of the research question.
\end{itemize}

\subsection*{4. Research Questions and Hypotheses}
\begin{itemize}
    \item Clearly state the primary research question(s) your group intends to answer.  
    \item Formulate the corresponding null and alternative hypotheses where applicable.
\end{itemize}

\subsection*{5. Data Source and Description}
\begin{itemize}
    \item Identify and describe the dataset(s) you will use. 
    \item Include information about the source, size, key variables, and how the data were collected.
    \item Include any known limitations or considerations with the dataset.
\end{itemize}

\subsection*{6. Preliminary Analysis Plan}
\begin{itemize}
    \item Outline the statistical methods and tests you plan to use (e.g., hypothesis tests, two-proportion tests).  
    \item Describe planned descriptive statistics, visualization techniques, and any data preprocessing steps.
\end{itemize}

\subsection*{7. Timeline and Roles}
\begin{itemize}
    \item Provide a rough schedule with milestones for your group’s work.  
    \item Assign preliminary tasks or responsibilities to group members.
\end{itemize}

\subsection*{Additional Notes}
\begin{itemize}
    \item If you have identified any challenges or questions so far, briefly list them here.
    \item Feel free to include references to sources or datasets if available.
\end{itemize}

\end{document}