\documentclass[12pt]{article}
\usepackage{resource/shortex}
\input{resource/teaching}

%--------------------
\begin{document}

\begin{center}
    \textbf{\Large Deliverable 2 \\ Final Project Progress Report}
\end{center}

You are required to submit a progress report slide for your final project along with the corresponding R code. 
This report should summarize the work completed so far, including data preparation, preliminary analysis, and any initial findings. 
Additionally, it should outline your plan for completing the project, covering remaining analyses and report writing.

\vspace{1em}

\subsection*{1. Project Title and Group Members}
\begin{itemize}
    \item Project Title
    \item Names of all Group Members
\end{itemize}

\subsection*{2. Summary of Project Goals}
\begin{itemize}
    \item Briefly restate the research questions and objectives of your project.
\end{itemize}

\subsection*{3. Data Description and Preparation}
\begin{itemize}
    \item Describe the dataset(s) you are using.
    \item Summarize any data cleaning or preprocessing steps completed so far.
    \item Discuss any issues or concerns encountered with the data (e.g., missing values, outliers).
\end{itemize}

\subsection*{4. Preliminary Analysis}
\begin{itemize}
    \item Present descriptive statistics and visualizations completed to date.
    \item Summarize any initial hypothesis testing, statistical analyses, or preliminary findings (if applicable).
    \item Include any key tables or figures, if appropriate.
\end{itemize}

\subsection*{5. Plan for Completion}
\begin{itemize}
    \item Outline your plan for completing the project, including remaining analyses, report writing, and division of tasks among group members.
    \item Provide an updated timeline or list of milestones.
\end{itemize}

\end{document}