---
title: "MA213 Basic Statistics and Probability - Lab1 Activity"
output:
  pdf_document: default
  html_document:
    code_download: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## **Lab 1: Introduction to R**

------------------------------------------------------------------------

## Lab Activities

#### Q1. Set up your working directory and import the file `baby_names.csv` into R. The data is on the blackboard and also it is available at:

#### <https://www.openintro.org/data/index.php?data=baby_names>

#### What does the dataset contain? Discuss with your partner. Consider the following:

-   How many rows and columns are in the dataset?

-   What kind of information does each column hold?

-   What kinds of questions or analyses could you perform with this data?

```{r, echo=TRUE}
#
#
#
```

#### Q2. Find the unique names in the `male` and `female` name columns. Assign them to variables `male_names` and `female_names`, and show how many unique names are in each column.

```{r, echo=TRUE}
#
#
#
```

#### Q3. Create a data frame from the table below. There are several ways to do this. There are several ways to do this; for example, you can build it column-wise or row-wise (using `append()` or `rbind()`).

| number | fruit  | size |
|--------|--------|------|
| 1      | apple  | 2    |
| 2      | apple  | 2    |
| 3      | apple  | 2    |
| 4      | orange | 4    |
| 5      | orange | 2    |
| 6      | orange | 2    |
| 7      | orange | 2    |
| 8      | orange | 3    |
| 9      | orange | 3    |
| 10     | orange | 3    |

```{r, echo=TRUE}
#
# 
#
```

#### Q4. Remove the `size` column from the data frame (assume the data frame is named `df`). Assign the new data frame without size to `df_2` and print it.

```{r, echo=TRUE}
#
#
#
```

#### Q5. Delete rows 9 and 10 from df. Assign the resulting data frame to `df_3` and print it.

```{r, echo=TRUE}
#
#
#
```

### Q6. Select all rows where fruit is "apple"

```{r, echo=TRUE}
#
#
#
```

### Q7. Select rows where fruit is "orange" and size equals 2

```{r, echo=TRUE}
#
#
#
```

### Q8. Select all rows where fruit is "apple"

```{r, echo=TRUE}
#
#
#
```

### Q9. Select rows where fruit is not "apple"

```{r, echo=TRUE}
#
#
#
```

#### Q10. Write all above R code to `lab1_lab.R` and submit it to GradeScope.

------------------------------------------------------------------------

## Post-lab activities

Please name your submission `Lab1.R` on GradeScope.

1.  Import `baby_names.csv` and assign it to the object `df_baby`.

2.  For the year 1999, what is the rank 1 boy’s name? Assign it to `boy_1999`.

3.  Add a registration number column to the dataset. The registration numbers should start from 1 and go up to 2820. Name this new column Registration (case-sensitive). Save the updated data frame as `new_df`.

Note: On GradeScope, the autograder will run your `lab1.R` to check that you have correctly created `df_baby`, `boy_1999`, and `new_df`.
