---
title: "MA213 Basic Statistics and Probability - Project 2 Lab (Week 2)"
bibliography: references.bib
output:
  pdf_document: default
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## Project 2 Lab — Week 2: **Working with Your Own Data**

------------------------------------------------------------------------

| Name | BUID | Present/Absent |
|------|------|----------------|
|      |      |                |
|      |      |                |
|      |      |                |
|      |      |                |

## Goals for Today

By the end of this lab you should have:

1.  Your dataset loaded and cleaned in R
2.  EDA completed for both your categorical and numerical variables
3.  Conditions for inference checked
4.  A first draft of your test statistic computed
5.  A clear plan for **Deliverable 2** (Results slides + R code + Datafile, due **April 7th**)

> **Reminder:** You may **not** use `t.test()`, `chisq.test()`, or `prop.test()`. All test statistics must be computed manually.\
> **ANOVA exception:** `aov()`, `anova()`, and `lm()` are permitted.

------------------------------------------------------------------------

## Part 1: Load and Inspect Your Data

### Step 1 — Load your dataset

Place your data file (`.csv`) in the same folder as this `.Rmd` file and set your working directory there.

```{r}
# Load your dataset
# Replace "your_data.csv" with your actual filename
df <- read.csv("your_data.csv")
```

```{r}
# How many rows and columns?
dim(df)
```

```{r}
# Variable names and types
str(df)
# or first few rows
head(df)
```

```{r}
# Quick summary of all variables
summary(df)
```

**Exercise 1.** Answer in the space below:

a.  How many observations (rows) and variables (columns) does your dataset have?

> **Answer:**

b.  List your two variables of interest — one categorical, one numerical. What data type does R assign to each (`character`, `factor`, `integer`, `numeric`)?

> **Categorical variable:**\
> **Numerical variable:**

c.  Are there any missing values (`NA`) in your key variables? Run the code below and report the result.

```{r}
# Check for missing values in your two key variables
# Replace var_cat and var_num with your actual variable names
# sum(is.na(df$var_cat))
# sum(is.na(df$var_num))
```

> **Answer:**

------------------------------------------------------------------------

## Part 2: Data Cleaning

Before doing any analysis, you need to make sure your data are in a usable form.

### Step 2 — Handle missing values and fix types

```{r}
# If your categorical variable is stored as character, convert to factor
# df$var_cat <- as.factor(df$var_cat)

# Drop rows with NA in your key variables if needed
# df <- df[!is.na(df$var_cat) & !is.na(df$var_num), ]

# Confirm cleaned dimensions
dim(df)
```

```{r}
# Check categories of your categorical variable
# unique(df$var_cat)
# table(df$var_cat)
```

**Exercise 2.**

a.  Did you need to remove any rows or fix any variable types? Describe what you did.

> **Answer:**

b.  How many categories does your categorical variable have? List them.

> **Answer:**

------------------------------------------------------------------------

## Part 3: Exploratory Data Analysis (EDA)

Good EDA is required in your Project 2 writeup. Complete all sections below for your own data.

### 3a — Categorical Variable EDA

**Exercise 3.** Create a frequency table and a bar plot for your categorical variable.

```{r}
# Frequency table
# tbl <- table(df$var_cat)
# tbl
```

```{r}
# Bar plot
# ggplot(df, aes(x = var_cat)) +
#   geom_bar(fill = "steelblue") +
#   labs(title = "Distribution of [Your Variable]",
#        x = "[Variable Label]", y = "Count") +
#   theme_minimal()
```

Describe what you see:

-   Which category has the highest / lowest count?
-   Is the distribution roughly equal, or highly skewed toward one category?

> **Answer:**

------------------------------------------------------------------------

### 3b — Numerical Variable EDA

**Exercise 4.** Compute summary statistics and create visualizations for your numerical variable.

```{r}
# Summary statistics
# mean_val   <- mean(df$var_num, na.rm = TRUE)
# median_val <- median(df$var_num, na.rm = TRUE)
# sd_val     <- sd(df$var_num, na.rm = TRUE)
# iqr_val    <- IQR(df$var_num, na.rm = TRUE)
# 
# cat("Mean:  ", round(mean_val, 3), "\n")
# cat("Median:", round(median_val, 3), "\n")
# cat("SD:    ", round(sd_val, 3), "\n")
# cat("IQR:   ", round(iqr_val, 3), "\n")
```

```{r}
# Histogram
# ggplot(df, aes(x = var_num)) +
#   geom_histogram(bins = 30, fill = "steelblue", color = "white") +
#   labs(title = "Distribution of [Your Variable]",
#        x = "[Variable Label]", y = "Count") +
#   theme_minimal()
```

```{r}
# If your analysis compares groups, add a boxplot by group
# ggplot(df, aes(x = var_cat, y = var_num, fill = var_cat)) +
#   geom_boxplot() +
#   labs(title = "[Numerical Variable] by [Categorical Variable]",
#        x = "[Group Variable]", y = "[Numerical Variable]") +
#   theme_minimal() +
#   theme(legend.position = "none")
```

Describe what you see:

-   Is the distribution symmetric or skewed?
-   Are there outliers?

> **Answer:**

------------------------------------------------------------------------

## Part 4: State Your Hypotheses

**Exercise 5.** Write out your hypotheses formally for both analyses. Use the space below.

### Categorical Analysis

What type of categorical test will you use? (Circle or bold one):

-   Inference for a single proportion (Ch. 6.1)
-   Inference for two proportions (Ch. 6.2)
-   Chi-square goodness-of-fit (Ch. 6.3)
-   Chi-square test for independence (Ch. 6.4)

$H_0$:

> 

$H_a$:

> 

------------------------------------------------------------------------

### Numerical Analysis

What type of numerical test will you use? (Circle or bold one):

-   One-sample t-test (Ch. 7.1)
-   Paired t-test (Ch. 7.2)
-   Two-sample t-test (Ch. 7.3)
-   ANOVA — comparing more than two groups (Ch. 7.5)

$H_0$:

> 

$H_a$:

> 

------------------------------------------------------------------------

## Part 5: Check Conditions for Inference

You must verify that conditions are met **before** computing any test statistic. This is a required section in your writeup.

**Exercise 6.** Check each condition below for **both** of your planned tests. Write "Met" or "Not met" and briefly explain your reasoning.

### Categorical Test Conditions

| Condition | Status | Explanation |
|----|----|----|
| Observations are independent |  |  |
| Data come from a random sample |  |  |
| Expected count $\geq 5$ in each cell |  |  |
| Sample size large enough ($n \geq 30$ or normal population) |  |  |

------------------------------------------------------------------------

### Numerical Test Conditions

| Condition                                               | Status | Explanation |
|---------------------------------------------------------|--------|-------------|
| Observations are independent                            |        |             |
| Data come from a random sample                          |        |             |
| Distribution is approximately normal OR $n \geq 30$     |        |             |
| Equal variance across groups (for ANOVA / two-sample t) |        |             |

```{r}
# Check normality visually with a QQ plot
# qqnorm(df$var_num)
# qqline(df$var_num, col = "red")
```

```{r}
# Check sample size per group
# table(df$var_cat)
```

> **Summary of condition checks:**

------------------------------------------------------------------------

## Part 6: First Draft of Your Test Statistic

Work through as much of the computation as you can today. It is fine if you do not finish — the goal is to make progress and identify where you are stuck.

### Categorical Test — First Draft

```{r}
# ── For chi-square goodness-of-fit ──────────────────────────────────────
# observed <- as.numeric(table(df$var_cat))
# n_total  <- sum(observed)
# k        <- length(observed)
# expected <- rep(n_total / k, k)        # adjust if H0 is not equal
#
# X2 <- sum((observed - expected)^2 / expected)
# df_cat <- k - 1
# p_val_cat <- 1 - pchisq(X2, df = df_cat)
#
# cat("Chi-square statistic:", round(X2, 4), "\n")
# cat("Degrees of freedom:  ", df_cat, "\n")
# cat("p-value:             ", round(p_val_cat, 6), "\n")

# ── For chi-square test of independence ─────────────────────────────────
# tbl      <- table(df$var_cat1, df$var_cat2)
# tbl_m    <- addmargins(tbl)
# r <- nrow(tbl); c_col <- ncol(tbl)
# total <- tbl_m[r+1, c_col+1]
#
# expected <- matrix(0, nrow = r, ncol = c_col)
# for (i in 1:r) {
#   for (j in 1:c_col) {
#     expected[i,j] <- tbl_m[i, c_col+1] * tbl_m[r+1, j] / total
#   }
# }
#
# X2 <- sum((tbl - expected)^2 / expected)
# df_cat <- (r - 1) * (c_col - 1)
# p_val_cat <- 1 - pchisq(X2, df = df_cat)
#
# cat("Chi-square statistic:", round(X2, 4), "\n")
# cat("Degrees of freedom:  ", df_cat, "\n")
# cat("p-value:             ", round(p_val_cat, 6), "\n")
```

------------------------------------------------------------------------

### Numerical Test — First Draft

```{r}
# ── For one-sample t-test ────────────────────────────────────────────────
# mu0    <- 0          # replace with your H0 value
# x_bar  <- mean(df$var_num, na.rm = TRUE)
# s      <- sd(df$var_num, na.rm = TRUE)
# n      <- sum(!is.na(df$var_num))
# SE     <- s / sqrt(n)
# t_stat <- (x_bar - mu0) / SE
# df_num <- n - 1
# p_val_num <- 2 * pt(-abs(t_stat), df = df_num)  # two-sided
#
# cat("t-statistic:", round(t_stat, 4), "\n")
# cat("df:         ", df_num, "\n")
# cat("p-value:    ", round(p_val_num, 6), "\n")

# ── For two-sample t-test ────────────────────────────────────────────────
# grp1 <- df$var_num[df$var_cat == "Group1"]
# grp2 <- df$var_num[df$var_cat == "Group2"]
#
# x1 <- mean(grp1); x2 <- mean(grp2)
# s1 <- sd(grp1);   s2 <- sd(grp2)
# n1 <- length(grp1); n2 <- length(grp2)
# SE <- sqrt(s1^2/n1 + s2^2/n2)
# t_stat <- (x1 - x2) / SE
# df_num <- min(n1, n2) - 1      # conservative df
# p_val_num <- 2 * pt(-abs(t_stat), df = df_num)
#
# cat("t-statistic:", round(t_stat, 4), "\n")
# cat("df:         ", df_num, "\n")
# cat("p-value:    ", round(p_val_num, 6), "\n")

# ── For ANOVA (aov() is allowed) ────────────────────────────────────────
# fit <- aov(var_num ~ var_cat, data = df)
# summary(fit)
```

**Exercise 7.** How far did you get? What is the one thing you need to figure out before next lab?

> **Answer:**

------------------------------------------------------------------------

## Part 7: Plan for Deliverable 2

**Deliverable 2** is due **April 7th** and must include:

-   Results slides (explaining your analysis)
-   R code file
-   Data file

**Exercise 8.** As a group, divide the remaining work using the table below. Fill it in together before you leave today.

| Task                           | Who is responsible | Target completion |
|--------------------------------|--------------------|-------------------|
| Final data cleaning            |                    |                   |
| Categorical EDA + inference    |                    |                   |
| Numerical EDA + inference      |                    |                   |
| Results slides (Deliverable 2) |                    |                   |
| Combine and review R code      |                    |                   |

> **Any questions or blockers to flag for your instructor:**
