---
title: "MA213 Lab Activity: Lab4 Activity"
subtitle: "Simulating Dice Games to Make Decisions"
output:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
```

# Introduction

You've been offered two different dice games. Before you decide whether to play, you need to understand:

1.  **Expected Value (E[X])**: What will happen *on average* in the long run?
2.  **Variance (Var[X])**: How much *variability* should you expect?

Today, you'll use both **theoretical calculations** and **simulations** to make informed decisions about these games!

------------------------------------------------------------------------

# Part 1: The Simple Game (One Die)

## Game Rules

> **You roll one die.**
>
> \- If you roll a **1** → You **WIN \$6**
>
> \- If you roll **2, 3, 4, 5, or 6** → You **LOSE \$1**
>
> **Question: Should you play this game?**

------------------------------------------------------------------------

## Step 1: Theoretical Expected Value

**Expected value** is the average outcome if you play the game many, many times.

**Formula:** $$E(X) = \sum_{i} x_i \cdot P(x_i)$$

Where $x_i$ are the possible outcomes and $P(x_i)$ are their probabilities.

### Calculate by hand first:

```{r theory-ev-simple}
# Possible outcomes and probabilities
outcomes <- c() # your turn; what are the outcomes?
probs <- c() # your turn; what are the corresponding probabilities?

# Expected Value
E_X <- sum(outcomes * probs)
E_X

cat("Expected Value:", E_X, "\n")
cat("Interpretation: On average, you gain $", round(E_X, 2), " per game.\n", sep="")
```

**Your calculation (show your work):**

E(X) =

**Decision based on Expected Value:**

\- If E(X) \> 0: **Play the game** (you gain money on average)

\- If E(X) \< 0: **Don't play** (you lose money on average)

\- If E(X) = 0: **Doesn't matter** (it's a fair game)

**Should you play?** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

------------------------------------------------------------------------

## Step 2: Theoretical Variance

**Variance** measures how much your winnings will *vary* from game to game.

**Formula:** $$Var(X) = E[(X - \mu)^2] = \sum_{i} (x_i - \mu)^2 \cdot P(x_i)$$

Or equivalently: $$Var(X) = E(X^2) - [E(X)]^2$$

```{r theory-var-simple}
# Method 1: Direct calculation
variance <- sum((outcomes - E_X)^2 * probs)
variance

# Method 2: E(X²) - [E(X)]²
E_X_squared <- NULL # your turn
variance_alt <- NULL  # your turn
variance_alt

# Standard deviation
sd_X <- sqrt(variance)

cat("Variance:", variance, "\n")
cat("Standard Deviation:", sd_X, "\n")
cat("\nInterpretation: Your winnings will typically vary by about $", 
    round(sd_X, 2), " from the expected value.\n", sep="")
```

**Your calculation (show your work):**

Var(X) =

**What does variance tell you?**

------------------------------------------------------------------------

## Step 3: Simulate the Game

Let's verify our theoretical calculations by **simulating** the game 100 times!

### Write a function for one game

```{r simulate-one-game}
# Function to simulate one game
play_simple_game <- function() {
  # your turn
  # output should be the dollar amount 
}

# Test it a few times
play_simple_game()
play_simple_game()
play_simple_game()
```

### Simulate 100 games

```{r simulate-many-games}
# Number of simulations
n_games <- 100

# Play the game n_games times
set.seed(123)  # For reproducibility
results <- replicate(n_games, play_simple_game())

# Calculate empirical statistics
empirical_mean <- mean(results)
empirical_var <- var(results)
empirical_sd <- sd(results)

# Compare with theoretical values
cat("=== COMPARISON: THEORETICAL vs SIMULATED ===\n\n")
cat("Expected Value:\n")
cat("  Theoretical: $", round(E_X, 4), "\n", sep="")
cat("  Simulated:   $", round(empirical_mean, 4), "\n", sep="")
cat("  Difference:  $", round(abs(E_X - empirical_mean), 4), "\n\n", sep="")

cat("Variance:\n")
cat("  Theoretical: ", round(variance, 4), "\n", sep="")
cat("  Simulated:   ", round(empirical_var, 4), "\n", sep="")
cat("  Difference:  ", round(abs(variance - empirical_var), 4), "\n\n", sep="")

cat("Standard Deviation:\n")
cat("  Theoretical: $", round(sd_X, 4), "\n", sep="")
cat("  Simulated:   $", round(empirical_sd, 4), "\n", sep="")
```

**Question:** Are the simulated values close to the theoretical values?

**Your answer:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Question:** What size n is needed for the simulated values to be close to the theoretical values? (increase `n_games`)

**Your answer:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

------------------------------------------------------------------------

## Step 4: Visualize the Results

### Cumulative Profit Over Time

```{r viz-cumulative}
# Calculate cumulative profit
n_games <- 100000

cumulative_data <- data.frame(
  game = 1:n_games,
  profit = cumsum(results),
  expected_line = (1:n_games) * E_X
)

# Plot cumulative profit
ggplot(cumulative_data, aes(x = game, y = profit)) +
  geom_line(color = "steelblue", alpha = 0.8) +
  geom_line(aes(y = expected_line), color = "red", linetype = "dashed", size = 1) +
  geom_hline(yintercept = 0, color = "black", linetype = "dotted") +
  annotate("text", x = 7500, y = max(cumulative_data$profit) * 0.8,
           label = "Expected trajectory", color = "red") +
  labs(title = "Cumulative Profit Over Time",
       subtitle = "Law of Large Numbers in Action!",
       x = "Number of Games Played",
       y = "Cumulative Profit ($)") +
  theme_minimal()
```

**Observation:**

------------------------------------------------------------------------

## Step 5: Understanding Variance Through Simulation

Let's see what variance really means by comparing results from different numbers of games.

```{r variance-understanding}
# Simulate different sample sizes
sample_sizes <- c(10, 50, 100, 500, 1000, 5000, 10000)
variance_data <- data.frame()

for (n in sample_sizes) {
  # Run 100 experiments of size n each
  experiments <- replicate(100, mean(replicate(n, play_simple_game())))
  
  variance_data <- rbind(variance_data, data.frame(
    n = n,
    mean_of_means = mean(experiments),
    sd_of_means = sd(experiments)
  ))
}

# Plot how variance of the mean decreases
ggplot(variance_data, aes(x = n, y = sd_of_means)) +
  geom_point(size = 3, color = "steelblue") +
  geom_line(color = "steelblue") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_x_log10() +
  labs(title = "How Uncertainty Decreases with More Games",
       subtitle = "Standard deviation of average winnings",
       x = "Number of Games Played (log scale)",
       y = "Standard Deviation of Mean Winnings") +
  theme_minimal()
```

**Key Insight:** What do you observe?

------------------------------------------------------------------------

# Part 2: The Two-Dice Game (Linear Combination)

## New Game Rules

> **You roll TWO dice.**
>
> **First die:**
>
> \- Roll a **1** → WIN \$6
>
> \- Roll **2-6** → LOSE \$1
>
> **Second die :**
>
> \- Roll a **1** → WIN \$3
>
> \- Roll **2-6** → LOSE \$3
>
> **(additional rule for the second die -\> double the outcome)**
>
> **Your total winnings = (First die result) + 2** $\times$ **(Second die result)**
>
> **Question: Should you play this new game?**

------------------------------------------------------------------------

## Step 1: Theoretical Expected Value (Two Dice)

Let $X_1$ = winnings from first die, $X_2$ = winnings from second die.

Your total winnings: $Y = X_1 + 2X_2$

**Property of Expected Value:** E(X_1 + 2X_2)=

```{r theory-ev-two-dice}
# First die (same as before)
E_X1 <- 6 * (1/6) + (-1) * (5/6)
cat("E(X1) = $", E_X1, "\n", sep="")

# Second die
outcomes_die2 <- c() # Your turn
probs_die2 <- c() # Your turn
E_X2 <- sum(outcomes_die2 * probs_die2)
cat("E(X2) = $", E_X2, "\n", sep="")

# Total expected value
E_Y <- c() # Your turn
cat("\nE(Y) = E(X1) + E(X2) = $", E_Y, "\n", sep="")
```

**Your calculation:**

$$E(X_1) $$

=

$$E(X_2) $$

=

$$E(Y) = E(X_1+2X_2) $$

=

**Should you play this game?** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

------------------------------------------------------------------------

## Step 2: Theoretical Variance (Two Dice)

**Property of Variance (for independent variables):** $$Var(X_1 + 2 X_2) $$

```{r theory-var-two-dice}
# Variance of first die (we already calculated this)
Var_X1 <- variance  # From Part 1
cat("Var(X1) =", Var_X1, "\n")

# Variance of second die
E_X2_squared <- sum(outcomes_die2^2 * probs_die2)
Var_X2 <- E_X2_squared - E_X2^2
cat("Var(X2) =", Var_X2, "\n")

# Total variance
Var_Y <- c() # Your turn
SD_Y <- c() # Your turn

cat("\nVar(Y) = Var(X1) + 4 * Var(X2) =", Var_Y, "\n")
cat("SD(Y) =", SD_Y, "\n")
```

**Your calculation:**

$$Var(X_1) $$ =

$$Var(X_2)$$

=

$$Var(Y) = Var(X_1 + 2X_2) $$

=

**Question:** Is the variance of the two-dice game larger or smaller than the one-die game?

**Your answer:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Why does this make sense?**

------------------------------------------------------------------------

## Step 3: Simulate the Two-Dice Game

```{r simulate-two-dice-game}
# Function to play the two-dice game
play_two_dice_game <- function() {
  # First die
  
  
  
  # Second die

  
  # Total winnings
  total <- c() # Your Turn : it should be the winning from the one play of two dice game 
  return(total)
}

# Test the function
play_two_dice_game()
play_two_dice_game()
play_two_dice_game()
```

### Run simulations

```{r simulate-two-dice-many}
# Simulate the game
n_games <- 10000 # change the number 
results_two_dice <- replicate(n_games, play_two_dice_game())

# Calculate empirical statistics
empirical_mean_2d <- mean(results_two_dice)
empirical_var_2d <- var(results_two_dice)
empirical_sd_2d <- sd(results_two_dice)

# Compare with theoretical values
cat("=== TWO-DICE GAME: THEORETICAL vs SIMULATED ===\n\n")
cat("Expected Value:\n")
cat("  Theoretical: $", round(E_Y, 4), "\n", sep="")
cat("  Simulated:   $", round(empirical_mean_2d, 4), "\n", sep="")
cat("  Difference:  $", round(abs(E_Y - empirical_mean_2d), 4), "\n\n", sep="")

cat("Variance:\n")
cat("  Theoretical: ", round(Var_Y, 4), "\n", sep="")
cat("  Simulated:   ", round(empirical_var_2d, 4), "\n", sep="")
cat("  Difference:  ", round(abs(Var_Y - empirical_var_2d), 4), "\n\n", sep="")

cat("Standard Deviation:\n")
cat("  Theoretical: $", round(SD_Y, 4), "\n", sep="")
cat("  Simulated:   $", round(empirical_sd_2d, 4), "\n", sep="")
```

**Question:** Do the simulated values match the theoretical values?

**Your answer:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

------------------------------------------------------------------------

## Step 4: Visualize Two-Dice Results

### Cumulative Profit Over Time

```{r viz-cumulative-two-dice}
# Calculate cumulative profit
cumulative_data_2d <- data.frame(
  game = 1:n_games,
  profit = cumsum(results_two_dice),
  expected_line = (1:n_games) * E_Y
)

# Plot
ggplot(cumulative_data_2d, aes(x = game, y = profit)) +
  geom_line(color = "coral", alpha = 0.8) +
  geom_line(aes(y = expected_line), color = "red", linetype = "dashed", size = 1) +
  geom_hline(yintercept = 0, color = "black", linetype = "dotted") +
  annotate("text", x = 7500, y = max(cumulative_data_2d$expected_line) * 0.8,
           label = "Expected trajectory", color = "red") +
  labs(title = "Cumulative Profit Over Time (Two-Dice Game)",
       subtitle = "Still converges to expected value!",
       x = "Number of Games Played",
       y = "Cumulative Profit ($)") +
  theme_minimal()
```

What do you observe?

**Your answer:**

------------------------------------------------------------------------

# Part 3: Reflection Questions

## Question 1: Expected Value

If both games have a **negative** expected value. What does this mean for you as a player?

**Your answer:**

------------------------------------------------------------------------

## Question 2: Variance and Risk

If the two-dice game has higher variance. What does this mean in practical terms?

**Your answer:**

------------------------------------------------------------------------

## Question 3:

In both simulations, we saw that as we played more games, the cumulative average got closer to the theoretical expected value.

a)  What does this tell you about short-term vs long-term gambling?

**Your answer:**

b)  If you only played 10 games, would you be guaranteed to end up at the expected value?

**Your answer:**

------------------------------------------------------------------------

## Question 4: Decision Making

Given what you learned:

a)  Would you play the one-die game? Why or why not?

**Your answer:**

b)  Would you play the two-dice game? Why or why not?

**Your answer:**

c)  If you had to choose between the two games, which would you choose?

**Your answer:**

------------------------------------------------------------------------

# Part 4: Your Own Game Design

**Challenge:** Design your own dice game!

Create a game where: 1. You have at least 2 different outcomes 2. Each outcome has a different payoff

Then: - Calculate the theoretical expected value - Calculate the theoretical variance - Simulate it 10,000 times - Compare theoretical vs simulated results - Visualize the distribution

```{r your-game-design}
# Your game design here

# 1. Define your game rules
# my_game <- function() {
#   # Your code here
# }

# 2. Calculate theoretical E(X) and Var(X)


# 3. Simulate 10,000 games


# 4. Compare theoretical vs simulated


# 5. Create visualizations


```

**Describe your game:**

**Theoretical expected value:**

**Theoretical variance:**

**Would you play your own game?**

------------------------------------------------------------------------

# Summary: Key Formulas

## Expected Value

**For a single random variable:** $$E(X) = \sum_{i} x_i \cdot P(x_i)$$

**For a linear combination:** $$E(aX + bY) = aE(X) + bE(Y)$$

## Variance

**For a single random variable:** $$Var(X) = E[(X - \mu)^2] = E(X^2) - [E(X)]^2$$

**For a linear combination (independent variables):** $$Var(aX + bY) = a^2Var(X) + b^2Var(Y)$$

## Standard Deviation

$$SD(X) = \sqrt{Var(X)}$$

------------------------------------------------------------------------

# Submission

Save your completed R Markdown file and submit The .Rmd file

Make sure all your answers are filled in and all code chunks run without errors!

**Student Name:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_ **BUID** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Student Name:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_ **BUID** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Student Name:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_ **BUID** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Date:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Did the simulations help you understand expected value and variance better?**

**Your reflection:**
