---
title: "MA213 Basic Statistics and Probability - Lab4 Guide"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## Lab 4: Simulation and Probability R Guide

------------------------------------------------------------------------

## Learning Objectives

- Validate and Explain Probability Distributions: Assess the validity of a probability distribution using the concepts of outcome, sample space, and probability properties (e.g., disjoint outcomes, probabilities between 0 and 1, and total probabilities summing to 1).

- Compute Probabilities Using Various Tools: Use logic, Venn diagrams, and probability rules to compute probabilities for events.

- Understand and Compute Expectations and Variances: Explain the concepts of expectations and variances of random variables, and compute the expectation and variance of a linear combination of random variables.

- Conduct Hypothesis Testing Using Simulation: Set up null and alternative hypotheses to test for independence between variables, and use simulation techniques to evaluate data support for these hypotheses.

------------------------------------------------------------------------

## How do you roll a die using R?

`sample()` will randomly sample `size` many samples from the vector `x`

`sample(x, size, replace = FALSE, prob = NULL)`

```{r sample-examples, echo=TRUE}
# Example
x <- c("apple", "pear", "strawberry", "orange", "lemon")

sample(x, 3) # picks 3 unique fruits from the vector (without replacement)
sample(x, 10, replace=TRUE) # picks 10 fruits, allowing repeats (sampling with replacement)
```

### Exercise 1: Make and roll a die once

**Task:** Create a vector for a fair six-sided die and roll it once.

```{r ex1-student, eval=FALSE}
# Your code here
die <- c(1, 2, 3, 4, 5, 6)
sample(die, 1)
```

```{r ex1-solution}
# Solution
die <- c(1, 2, 3, 4, 5, 6)
# Or simply:
die <- 1:6

# Roll once
sample(die, 1)
```

### Exercise 1b: Roll a weighted die

**Task:** Create a six-sided die where 6 appears twice as often as other numbers. Roll it 20 times.

```{r ex1b-student, eval=FALSE}
# Your code here
#
#
```

```{r ex1b-solution}
# Solution
die <- 1:6
# Probabilities: 6 has twice the probability of others
probs <- c(1/7, 1/7, 1/7, 1/7, 1/7, 2/7)

# Roll 20 times
sample(die, 20, replace = TRUE, prob = probs)

# Verify the probabilities work
sum(probs)  # Should equal 1
```

------------------------------------------------------------------------

## Loops

Loops repeatedly execute a block of code for various elements.

Example: Print numbers from 1 to 5

```{r loop-example1, echo=TRUE}
# Example 1: Print numbers from 1 to 5
print("Example 1")
for (i in 1:5) {
  print(i)
}
```

### Example 2: Print numbers from 1 to 5

```{r loop-example2, echo=TRUE}
print("Example 2")
for (i in c(1, 2, 3, 4, 5)) {
  print(i)
}
```

### Example 3: Simulate rolling a die 5 times and record it

```{r loop-example3, echo=TRUE}
print("Example 3")
rolls <- rep(0, 5) # Initialize a vector of length 5 with zeros to store roll results
for (i in 1:5) {
  rolls[i] <- sample(1:6, 1)   # Sample one number between 1 and 6
}
rolls
```

### For loop syntax

``` r
# For loop syntax
for (variable in sequence) {
    # code you want to execute repeatedly until the end of the sequence     
}
```

### Exercise 2: Make a coin and flip it 10 times

**Task:** Create a coin (a vector of size 2) and flip it 10 times. Record each flip result to a vector `coin_exp`.

```{r ex2-student, eval=FALSE}
# Your code here
x <- c(0, 1) # here is a coin. 0 for tail and 1 for head

#
#
```

```{r ex2-solution}
# Solution
x <- c(0, 1) # 0 for tail, 1 for head
coin_exp <- rep(0, 10) # Initialize vector

for (i in 1:10) {
  coin_exp[i] <- sample(x, 1)
}

coin_exp

# Alternative solution using sample directly
coin_exp_alt <- sample(c(0, 1), 10, replace = TRUE)
coin_exp_alt
```

### Exercise 2b: Flip a biased coin

**Task:** Flip a coin 100 times where heads (1) appears 70% of the time. Calculate the proportion of heads.

```{r ex2b-student, eval=FALSE}
# Your code here
#
#
```

```{r ex2b-solution}
# Solution
coin <- c(0, 1)
probs <- c(0.3, 0.7) # 30% tails, 70% heads

flips <- sample(coin, 100, replace = TRUE, prob = probs)

# Calculate proportion of heads
prop_heads <- mean(flips)
prop_heads

# Visualize
table(flips)
barplot(table(flips), 
        main = "Biased Coin Flips (100 flips)",
        xlab = "Outcome (0=Tail, 1=Head)",
        ylab = "Frequency",
        col = c("coral", "steelblue"))
```

------------------------------------------------------------------------

## Functions

### Function syntax

``` r
my_function <- function(arg1, arg2 = default_value) {
  # code to execute
  result <- arg1 + arg2    # example operation
  return(result)           # optional; last expression is returned anyway
}
```

Minimal form (relying on implicit return):

``` r
add <- function(x, y) {
  x + y   # value of last expression is returned
}
add(3, 4)
```

Function with no arguments:

``` r
greet <- function() {
  cat("Hello, world!\n")
}
greet()
```

### Example: Function to calculate (a+b)\^2

```{r function-example, echo=TRUE}
# Example: Function to calculate (a+b)^2
square_sum <- function(a, b) {
  result <- (a + b)^2
  return(result)
}

square_sum(3, 4)
square_sum(10, 5)
```

### Exercise 3: Multiplication of three numbers

**Task:** Make a function to calculate the multiplication of three numbers.

```{r ex3-student, eval=FALSE}
# Your code here
# Think about function name
# Think about inputs and outputs

#
#
```

```{r ex3-solution}
# Solution
multiply_three <- function(x, y, z) {
  x * y * z
}

# Test
multiply_three(2, 3, 4)
multiply_three(5, 10, 2)
```

### Exercise 3b: Mean and standard deviation function

**Task:** Create a function that takes a vector of numbers and returns both the mean and standard deviation.

```{r ex3b-student, eval=FALSE}
# Your code here
#
#
```

```{r ex3b-solution}
# Solution
stats_summary <- function(x) {
  mean_val <- mean(x)
  sd_val <- sd(x)
  
  list(mean = mean_val, sd = sd_val)
}

# Test
test_data <- c(2, 4, 6, 8, 10, 12)
stats_summary(test_data)

# Alternative: using return explicitly
stats_summary2 <- function(x) {
  return(list(
    mean = mean(x),
    sd = sd(x),
    n = length(x)
  ))
}

stats_summary2(test_data)
```

### Exercise 3c: Temperature converter

**Task:** Create a function that converts Fahrenheit to Celsius. Formula: C = (F - 32) × 5/9

```{r ex3c-student, eval=FALSE}
# Your code here
#
#
```

```{r ex3c-solution}
# Solution
fahrenheit_to_celsius <- function(fahrenheit) {
  celsius <- (fahrenheit - 32) * 5/9
  return(celsius)
}

# Test
fahrenheit_to_celsius(32)   # Should be 0
fahrenheit_to_celsius(212)  # Should be 100
fahrenheit_to_celsius(98.6) # Should be 37

# Bonus: Create vector conversion
temps_f <- c(32, 50, 68, 86, 104)
temps_c <- fahrenheit_to_celsius(temps_f)
data.frame(Fahrenheit = temps_f, Celsius = temps_c)
```

------------------------------------------------------------------------

## If Statements (Conditional Logic)

If statements let your code make decisions based on conditions.

### Basic Syntax

```{r if-syntax, echo=TRUE}
# Simple if
x <- 5
if (x > 0) {
  print("Positive")
}

# If-else
number <- 7
if (number %% 2 == 0) {
  print("Even")
} else {
  print("Odd")
}

# If-else if-else
score <- 85
if (score >= 90) {
  grade <- "A"
} else if (score >= 80) {
  grade <- "B"
} else {
  grade <- "C"
}
print(grade)
```

### Comparison Operators

```{r operators, echo=TRUE}
# ==  equal to
# !=  not equal to
# >   greater than
# <   less than
# >=  greater than or equal
# <=  less than or equal
# %in% check if value in a set

5 == 5        # TRUE
5 != 4        # TRUE
5 > 3         # TRUE
3 %in% c(1,2,3)  # TRUE
```

### Logical Operators

```{r logical, echo=TRUE}
# & (AND) - both must be TRUE
(5 > 3) & (10 > 8)   # TRUE

# || (OR) - at least one must be TRUE  
(5 > 3) || (10 < 8)   # TRUE

# ! (NOT) - reverses the condition
!(5 < 3)              # TRUE
```

### Example: Dice Game with If Statements

```{r dice-if-example, echo=TRUE}
play_dice <- function() {
  roll <- sample(1:6, 1)
  
  if (roll == 1) {
    winnings <- 10
  } else if (roll >= 2 && roll <= 4) {
    winnings <- 1
  } else {
    winnings <- -2
  }
  
  cat("Rolled:", roll, "| Won: $", winnings, "\n")
  return(winnings)
}

play_dice()
play_dice()

# Expected value
# P(1)=1/6→$10, P(2-4)=3/6→$1, P(5-6)=2/6→-$2
ev <- (1/6)*10 + (3/6)*1 + (2/6)*(-2)
cat("Expected value:", ev)
```

### Exercise: Two-Coin Game

**Task:** Flip two coins. Both heads: win \$5, both tails: lose \$3, one each: \$0

```{r coin-exercise, eval=FALSE}
# Your code here
coin_game <- function() {
  coin1 <- sample(c(0, 1), 1)
  coin2 <- sample(c(0, 1), 1)
  
  # Use if statements to determine winnings
  
}
```

```{r coin-solution}
# Solution
coin_game <- function() {
  coin1 <- sample(c(0, 1), 1)
  coin2 <- sample(c(0, 1), 1)
  
  if (coin1 == 1 && coin2 == 1) {
    winnings <- 5
  } else if (coin1 == 0 && coin2 == 0) {
    winnings <- -3
  } else {
    winnings <- 0
  }
  
  return(winnings)
}

# Test
coin_game()

# Expected value: P(HH)=1/4→$5, P(TT)=1/4→-$3, P(HT or TH)=1/2→$0
ev <- (1/4)*5 + (1/4)*(-3) + (1/2)*0
cat("Expected value:", ev)
```

**Key Points:** - Use `==` for comparison (not `=`) - Use `&&` for AND, `||` for OR in if statements - Always use curly braces `{}`

------------------------------------------------------------------------

## Combining Functions and Loops

You can call loops inside functions for repeated simulations.

Example: Roll a die multiple times

```{r loop-function-example, echo=TRUE}
# Example: Roll a die 5 times using a loop within a function
roll_multiple <- function(runs) {
  rolls <- rep(0, runs) # create rolls vector of 0
 
  for (i in 1:runs) {
    rolls[i] <- sample(1:6, 1)
  }
  
  return(rolls)
}

roll_multiple(5)
```

### Exercise 4: Coin flip function

**Task:** Create a function that flips a coin n times and returns the proportion of heads.

```{r ex4-student, eval=FALSE}
# Your code here
#
#
```

```{r ex4-solution}
# Solution
flip_coin_prop <- function(n) {
  coin <- c(0, 1) # 0 = tails, 1 = heads
  flips <- rep(0, n)
  
  for (i in 1:n) {
    flips[i] <- sample(coin, 1)
  }
  
  proportion_heads <- mean(flips)
  return(proportion_heads)
}

# Test
flip_coin_prop(10)
flip_coin_prop(100)
flip_coin_prop(1000)
```

### Exercise 4b: Compare different sample sizes

**Task:** Use the function from Exercise 4 to compare proportions from 10, 100, 1000, and 10000 flips.

```{r ex4b-student, eval=FALSE}
# Your code here
#
#
```

```{r ex4b-solution}
# Solution
sample_sizes <- c(10, 100, 1000, 10000)
proportions <- numeric(length(sample_sizes))

for (i in seq_along(sample_sizes)) {
  proportions[i] <- flip_coin_prop(sample_sizes[i])
}

results <- data.frame(
  Sample_Size = sample_sizes,
  Proportion_Heads = proportions
)
results

# Visualize convergence to 0.5
ggplot(results, aes(x = Sample_Size, y = Proportion_Heads)) +
  geom_point(size = 3, color = "steelblue") +
  geom_line(color = "steelblue") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  scale_x_log10() +
  labs(title = "Law of Large Numbers: Coin Flips",
       x = "Number of Flips (log scale)",
       y = "Proportion of Heads") +
  theme_minimal()
```

------------------------------------------------------------------------

## Two-Dice Example

Roll two dice once and compute their sum:

```{r two-dice-once}
# Outcome from those two dice
dice1 <- sample(1:6, 1)
dice2 <- sample(1:6, 1)

sum_of_two <- dice1 + dice2
sum_of_two
```

Simulate two-dice sum for 100 iterations using a for loop:

```{r two-dice-loop}
myoutcome <- integer(100)
for (i in 1:100) {
  dice1 <- sample(1:6, 1)
  dice2 <- sample(1:6, 1)
  myoutcome[i] <- dice1 + dice2
}
head(myoutcome)
```

### Using replicate()

`replicate()` performs repeated runs efficiently and succinctly.

Define one simulation run:

```{r replicate-function}
twodice_outcome <- function() {
  sum(sample(1:6, 2, replace = TRUE))  # sample two at once and sum
}
twodice_outcome()
```

Run many simulations:

```{r replicate-run}
myoutcome <- replicate(n = 30, twodice_outcome())
myoutcome
```

### Exercise 5: Three-dice simulation

**Task:** Create a function to roll three dice and return their sum. Simulate 1000 rolls.

```{r ex5-student, eval=FALSE}
# Your code here
#
#
```

```{r ex5-solution}
# Solution
threedice_outcome <- function() {
  sum(sample(1:6, 3, replace = TRUE))
}

# Simulate 1000 rolls
outcomes_3dice <- replicate(1000, threedice_outcome())

# Summary statistics
mean(outcomes_3dice)
sd(outcomes_3dice)

# Visualize
hist(outcomes_3dice, 
     breaks = 15,
     main = "Distribution of Three Dice Sum",
     xlab = "Sum of Three Dice",
     col = "lightblue",
     border = "white")
```
