\documentclass[12pt]{article}
\usepackage{resource/shortex}
\input{resource/teaching}


%--------------------
\begin{document}

\begin{center}
\textbf{\Large Project 1 Guide \\ Week 7 } 
\end{center}

\section*{Pre-Lab Activities}
Let's review the project 1 requirements. 
Think about your role and responsibilities in your group.
Please go over the following questions and be ready to discuss them in the lab session.

\section*{Lab Activities}

At this point, your group will go over the final checklist for the project 1 video presentation. Let's discuss whether you have completed the following sections.

\begin{enumerate}
  \item Does the introduction part convey motivation, data, results, and takeaways?
  
  \vspace{1\baselineskip}
  
  \item Does your group include both numerical data analysis and categorical data analysis?  
  Does the categorical data analysis part include any contingency table and bar plot to discuss the relationship between two categorical variables?  
  Does the numerical data analysis part include summary statistics and visualizations (e.g., histogram, box plot) to describe the distribution of a numerical variable?
  
  \vspace{1\baselineskip}
  
  \item Does the conclusion summarize key findings and takeaways from both numerical and categorical data analyses?
  
  \vspace{1\baselineskip}
  
  \item Does your group video presentation satisfy the technical requirements (e.g., length, format, etc.)?
\end{enumerate}

\section*{Post Lab Activities}

You have a project deliverable: the project 1 video due next Friday at 8:00 AM.



\end{document}
