\documentclass[12pt]{article}
\usepackage{resource/shortex}
\input{resource/teaching}
\usepackage[hidelinks]{hyperref}
\setlength{\parskip}{0.5em}

%--------------------
\begin{document}

\begin{center}
\textbf{\Large Project 2 Guide \\ Week 10 } 
\end{center}

\section*{Lab Activities}

\subsection*{1. Check-in and Progress Sharing}
Each group member briefly shares progress since the last lab:  
\begin{itemize}
    \item Data collection and cleaning status  
    \item Preliminary analysis and visualizations completed  
    \item Any interesting patterns or initial findings  
    \item Problems or challenges encountered  
\end{itemize}

\vspace{2\baselineskip}

\subsection*{2. Review and Discussion of Preliminary Results}
Discuss your group's current results and analyses:  
\begin{itemize}
    \item Are the descriptive statistics and visualizations clear and informative?  
    \item Do the initial hypothesis tests make sense?  
    \item Is the data quality sufficient for your analysis?  
    \item Identify any areas that require re-analysis or additional checks.  
\end{itemize}

\vspace{2\baselineskip}

\subsection*{3. Progress Report Preparation}
Begin drafting key sections of your progress report:  
\begin{itemize}
    \item Revisit and refine your research questions and hypotheses if needed.  
    \item Summarize the dataset(s) and data preparation steps completed to date.  
    \item Include descriptive statistics and preliminary analysis results (tables, graphs).  
    \item Identify outstanding challenges and questions for instructors/TAs.  
    \item Outline your updated plan and timeline for completing the final project.  
\end{itemize}

\vspace{2\baselineskip}

\subsection*{4. Group Roles and Collaboration Check}
Review the assigned roles and responsibilities:  
\begin{itemize}
    \item Confirm each member’s current contributions and tasks.  
    \item Adjust roles if needed to address current project needs.  
    \item Discuss communication and file sharing best practices for smooth collaboration.  
\end{itemize}

\vspace{2\baselineskip}


\section*{Post Lab Activities}

\begin{itemize}
    \item Please complete and submit your \textbf{Progress Report} by 8:00 AM next Friday.
    \item Begin drafting sections of the final report based on progress so far.  
    \item Schedule your next group meeting(s) to finalize analysis and report writing.  
\end{itemize}

\section*{Reminders}

\begin{itemize}
    \item Ensure your progress report includes properly labeled figures and tables.  
    \item Keep track of all data sources and references for your final report.  
    \item Contact the instructor, TF or LA well in advance if you anticipate difficulties meeting the deadline.  
\end{itemize}

\end{document}