---
title: "MA213 Basic Statistics and Probability - Bonus Lab Guide"
output:
  html_document:
    code_download: true 
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Post Lab Activity Bonus Lab

------------------------------------------------------------------------

1.  Contingency Table using `dplyr`

A contingency table displays the frequency distribution of variables. You can create contingency tables using `dplyr` in combination with `tidyr`.

For example, to create a contingency table of Month vs. Ozone_Level:

```{r}
df <- read.csv("airquality.csv")

# filter out all rows with any missing values
filtered_df <- df %>%
  filter(complete.cases(.))

# classify each observation as having "High" or "Low" ozone depending on whether its Ozone value exceeds the median (ignoring NAs)
filtered_df2 <- filtered_df %>%
  mutate(Ozone_Level = ifelse(Ozone > median(Ozone, na.rm=TRUE), "High", "Low"))

ozone_median <- median(filtered_df$Ozone, na.rm = TRUE)

# equivalent code
filtered_df2 <- filtered_df %>%
  mutate(
    Ozone_Level = case_when(
      Ozone > ozone_median ~ "High",
      Ozone > 0 & Ozone <= ozone_median  ~ "Low"
    )
  )


# create a contingency table that shows the count of "High" and "Low" ozone observations for each month.
contingency_table <- filtered_df2 %>%
  group_by(Month, Ozone_Level) %>%
  summarise(Count = n()) %>%
  pivot_wider(names_from = Ozone_Level, values_from = Count)


print(contingency_table)
```

2.  Show histogram of `Ozone`.

```{r}

Hist_Ozone<-ggplot(data=df)+geom_histogram(aes(Ozone))

print(Hist_Ozone)
```

3.  You have a slightly biased coin where $P(head)=0.40$ and $P(tail)=0.60$.

You flip the coin twice. If the first outcome is head (that means 1) you earn 2 dollars and tail (that means 0) for nothing. If the second outcome is head, you need to pay a dollar and tail for nothing.

We are working on this simulation procedure. We are creating a function called `one_trial()` which takes no input values (it is fine for the function to have no inputs) and outputs the amount of money earned from one trial of the game (the money obtained from flipping twice).

``` r

one_trial <- function(){

  first <- sample(c(1,0), 1, prob=c(0.4, 0.6)) # output from the first flip
  second <- sample(c(1,0), 1, prob=c(0.4, 0.6))  # output from the second flip  
  out <- first + second # This line needs to be modified ... 
  return(out)
}
one_trial()
```

4.  Recent statistics suggest that 10% of Americans rely on Social Network Sites (SNSs). We are going to have a simulation study if we can verify that CLT really works in this case too. Let's assume that US population is 300,000 for calculation purposes.

    Generate the population data and assign it to a variable called `population`. This `population` should consist of 300,000 values, each being either `1` or `0` where `1` indicates reliance on SNSs and `0` indicates not relying on SNSs.

```{r}
population <- c(rep(1, 30000), rep(0, 270000))
```

5.  Testing for independence in two-way tables.

```{r}
    # loading from csv file
    df <- read.csv("titanic.csv")

    df2 <- df[,2:3]
    head(df2)

    unique(df2$Survived)
    unique(df2$Pclass)

    tbl <- table(df2)
    tbl

    tbl2 <- addmargins(tbl)
    tbl2

    col1sum <- tbl2[3]
    col2sum <- tbl2[6]
    col3sum <- tbl2[9]
    row1sum <- tbl2[10]
    row2sum <- tbl2[11]
    total <- tbl2[12]


    # Calculate expected counts for each cell
    e11 = row1sum * col1sum / total
    e12 = row1sum * col2sum / total
    e13 = row1sum * col3sum / total
    e21 = row2sum * col1sum / total
    e22 = row2sum * col2sum / total
    e23 = row2sum * col3sum / total

    # Create 2x3 matrix of expected counts, filled by row
    expected_tbl = matrix(data = c(e11, e21, e12, e22, e13, e23),
                          nrow = 2, ncol = 3) 

    # Print the expected table
    expected_tbl

    # Compute the chi-square statistic
    statistic <- 0
    for (i in 1:6){
      statistic <- statistic + (tbl[i] - expected_tbl[i])^2/expected_tbl[i]
    }
    statistic


    # p-value
    df = (2-1)*(3-1)
    df
    pvalue = 1-pchisq(statistic, df)
```

6.  Obtain b0 and b1 from the `Beauty in the class` data using the model $$
    y = \beta_0 + \beta_1 x + \epsilon
    $$ where b0 and b1 are the estimates of the intercept ($\beta_0$) and slope ($\beta_1$) parameters in the linear regression relating $x$ (beauty score) to $y$ (course evaluation).

```{r}

    df <- read.csv("beauty.csv")
    model = lm(eval~beauty, data=df)
    model_summary <-summary(model)
    model_summary$coefficients
    b0 <- model_summary$coefficients[1]
    b1 <- model_summary$coefficients[2]

    b0
    b1
```
